A Quantitative Genetic Linked Selection Model fit to the Negative Selection Process in Humans

### 2023-07-06 manuscript/manuscript.tex
ms changes

\subsection*{Predicted Diversity, Substitution Rates, and Fitness Variance}

@@

When we fit our negative selection model to genome-wide patterns of diversity,
the parameter estimates generate three predicted quantities: (1) the $B(x)$
map, (2) the substitution rates per segment, and (3) the fitness variance per
segment. Here we focus on the predicted $B(x)$ map and implied levels of
diversity at the megabase scale. We discuss substitution rates and fitness
variance in the next section.

We also obtain a prediction $\widehat{B}_i$ for the reduction factor $B_i$ in
each window $i$, which can be multiplied by our estimate of $\widehat{\pi}_0$
to obtain a prediction of diversity $\widehat{\pi}_i = \widehat{\pi}_0
\widehat{B}_i$.

@@

@@

\subsection*{Evolutionary Models}

We model the diversity at punitively neutral sites. These are sites in regions
we have \emph{a prior} reason to believe have a DFE nearly entirely
concentrated in the neutral range. While this may not be the case in reality,
it is necessary given the computational demands of modeling whole-genome
selection.

CDS+phastcons+UTR model

 - UTR as neutral control

 - CDS takes priority over phastcons, phastcons is fallback. Otherwise, the
   effect of CDS would be biased by the fraction removed and classified as
   phastcons. This is the right comparison for the substitution model.

\subsection*{Predicted Substitution Rates}

- mention neg. auto correlation?

- Gene level analysis

To evaluate the predicted substitution rate of our CDS+phastcons+UTR model,

Additionally, our GBGS model makes predictions about the rate of substitutions
in different segments under purifying selection. There are two stages of the
prediction process. First, when the initial B' maps are calculated, the
predicted substitution rate is calculated as a byproduct of

@@

% As genomes are transmitted through the generations, lifetime events selectively
% alter what is passed to the next generation. While selection directly acts on
% the genetic variants that cause heritable fitness differences among
% individuals, neighboring regions experience an indirect contraction in genetic
% variability. This byproduct is known as \emph{linked selection}, and it occurs
% when selection changes a site's frequency on timescale faster than
% recombination can decouple that site from its genetic background. Because
% linked selection modifies the genetic variability in a regionally-correlated
% way, it creates observable patterns in genomic data that are used to understand
% the type of selection processes acting on genomes and estimate key evolutionary
% quantities.

% , with characteristic length
% $\nicefrac{s}{r}$ basepairs, where $s$ is the heterozygous selection
% coefficient and $r$ is the per-basepair recombination rate. Sweeps were found
% in humans, \emph{Drosophila}, and other species (XXX), though it was unclear
% whether hitchhiking alone could explain the observed genome-wide correlation
% between recombination rate and genetic diversity in some species (XXX).

% Much work has been done on developing mathematical theory for two linked
% selection processes: hitchhiking and background selection. These are
% distinguished by the evolutionary processes that generate fitness variation and
% the resulting selection dynamics. Under the classic hitchhiking model, a new
% beneficial mutation sweeps fixation, leading to a dip in genetic diversity on a
% small spatial scale. Since the sweeps leave relatively narrow footprint, they
% influence genome-wide diversity only when beneficial mutations are recurring
% and recombination is relatively limited, as in \emph{Drosophila}. These models
% assume sweeps are sufficiently rare as to not occur simultaneously, which is
% consistent with the slow estimated rate of strongly beneficial mutations (XXX).
% Overall, the classic hitchhiking model is simple because the underlying
% selection process does not generate polygenic fitness variation, even though
% this is type of variation plays a dominant role in adaptation (XXX).

% By contrast, background selection is a consequence of the fitness variation
% created by a steady influx of new deleterious mutations. The classic background
% selection model (XXX) considers how the balance between mutation introducing
% deleterious variants into the population and selection removing them, reduces
% nearby genetic variability. A critical assumption of this model is that such an
% equilibrium between mutation and selection exists, which is the case only when
% these mutations have a large enough fitness cost to ensure that they cannot fix
% in the population. Under this process, the number of deleterious mutations per
% gamete reaches a Poisson stationary distribution with rate $\nicefrac{U}{s}$,
% where $U = \mu_D L$ is the deleterious mutation rate ($\mu_D$ per basepair per
% generation) in a region of $L$ basepairs for mutations with a heterozygous
% selection coefficient of $s$ (XXX). The fraction of gametes without any
% mutations is then $f_0 = \exp(-\nicefrac{\mu_D}{s})$. Without recombination,
% any gamete with deleterious mutations is destined to bring its linked neutral
% variation with it to extinction, effectively reducing diversity levels to those
% expected in a neutrally-evolving population of size $f_0 N$.

% Remarkably, predictions from simple background selection models with
% recombination closely match large-scale patterns of human diversity and
% divergence (XXX). Since background approximately rescales a neutral coalescent
% process by $f_0$, the predicted diversity is reduced from neutral levels by a
% factor $B = \nicefrac{N_e}{N}$, which is the ratio of effective population size
% under background selection to the population size. In background selection
% models with recombination, $B$ is determined by the number of conserved
% basepairs and recombination in a region. Consequently the predicted reduction
% along the genome (known as $B$ maps) mirror the regional density of conserved
% elements under purifying selection and recombination rates. These $B$ maps have
% been used in studies of XXX, are inputs into machine learning models that
% predict functional constraint, and XXX. Furthermore, since new deleterious
% mutations are perpetually flowing into populations $B$ maps may provide a
% better null model for evolution than the standard neutral model.

% While background selection models do predict diversity levels  XXX

% % outline?
% - Disentangling evolutionary processes from the patterns of genetic variation
% across the genome. The structure of what can be under selection shapes these
% patterns when selection acts; when drift acts, these patterns should not exist.

% - For species with small effective population sizes (and perhaps large map
% lengths), BGS seems to be a dominant selective force in shaping patterns of
% diversity (e.g. at the megabase scale in humans).

% - However, current approaches to fitting BGS models fail under the weak
% selection domain. This weak selection domain is important (discuss
% HRI/ratchet/etc). Importantly, the characteristic scale of the impact of linked
% selection is $\nicefrac{s}{r}$, meaning weak selection acts on smaller scales
% than strong selection. Our ability to see selection at small scales may be
% inhibited by the fact that our current BGS theory only works for strong
% selection.

% - We need to say briefly that we're using BGS to mean linked selection due to
% purifying selection, and thus the strong and weak regimes (classically it has
% just meant the strong regime).

% \section*{Conclusion}

% - care should be taken applying these methods to \emph{Drosophila}, since the
% $\delta_{LD}$ term could be much larger.

% \section*{Results}

% \subsection*{Theory}

% - overview of why BGS models fail?

% From an evolutionary perspective, linked selection is important because by
% modifying neighboring levels of genetic diversity, it is also directly altering
% the selection dynamics at these neighboring sites.

% The other use of these models is to understand how linked selection processes
% differentially alter the genetic variability, or equivalently, the genealogical
% trees, along the genome.

% XXX how many sites are under selection at any moment? What's fitness variation?

% Early on, some pointed out that evidence of strong selective sweeps in the
% human genome was not the full story

% Generally, in organisms with large amounts of recombination, linked selection
% has

% Weak selection processes

% distorts
% genome-wide

% diversity.

% The linked selection literature has primarily focused on developing theory for,
% and quantifying the effects of two linked selection processes: hitchhiking and
% background selection. Genome-wide models of the hitchhiking process consider
% how beneficial mutations that sporadically enter the population

% - QGLS polygenic linked selection

% - H and W

% - inflation factor

% \subsection*{Statistical Methods}

TODO we need to make appoint about how Ben G's paper found the V(n)->VA connectoin -- this is their patching model

TODO we need to give some intuition for what Vm and V are in humans, e.g. back of envelope type stuff

% The $N_{e,t}$ terms are analogous to an instantaneous coalesce rate at
% generation $t$, which can be expressed in terms of the variance in allele
% frequency change by noting that in an ideal Wright--Fisher population,

% \begin{align}
%   \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N}\right)^t \right]
% \end{align}

% This is an asymptotic $N_e$, but this does not accurately describe the levels
% of diversity (\cite{Santiago1998-bs} equation 16, \cite{Santiago2016-mu}
% equation 4).

% \begin{align}
%   \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right) + \sum_{i \ne j} C_{i,j}
% \end{align}

% In general, the variance in allele frequency after $t$ generations
% of neutral evolution is,

%\begin{align}
%  \label{eq:var_freq}
%  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
%  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
%\end{align}

% Each term is time-indexed because it is conditioned on the starting neutral
% frequency in each individual, $x_{0,1}$. This equation is very general, and
% allows for any reproduction model with finite offspring variance and can
% account for the impact of linked selection under a variety of selection
% processes.

% Next, we can partition the gamete number into heritable ($f_i$) and
% non-heritable ($\varepsilon_i$) fitness components of each individual $i$, $k_i
% = f_i + \varepsilon_i$. We define $f_i$ such that $\nicefrac{1}{N} \sum_i f_i =
% 2$ in a constant population, and $\varepsilon_i$ is noise term centered around
% zero. Then, the neutral allele frequency change is can be written as the sum of
% three independent stochastic terms,

% \begin{align}
%   \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
%                \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
%   + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
%   % \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
%                % \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
%   % + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
%   \Delta p_1 &= S_1 + D_1 + H_1
% \end{align}

% (c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).

% Each term is time-indexed because it is conditioned on the starting neutral
% frequency in each individual, $x_{0,1}$. This equation is very general, and
% allows for any reproduction model with finite offspring variance and can
% account for the impact of linked selection under a variety of selection
% processes.

% Now, let us look the variance of $\var(\Delta p_1)$ over evolutionary
% replicates.

% \begin{align}
%   \var(\Delta p_1) &= \var(S_1) + \var(D_1) + \var(H_1)
% \end{align}

% To introduce how this equation can be used to quantify the variance in allele
% frequency change through time, $\var(p_t - p_0$ for $t > 0$, we first look at
% evolution in a neutral Wright--Fisher model. Neutrality ensures that $S_t = 0$
% for all $t$.

%%% OLD

%There has been considerable past work to find both the rate of the ratchet
%(XXX), and summaries of the genealogies created under weak purifying selection
%(XXX). Our approximation combines a few theoretic avenues from past work. Our
%primary goal is to find a theoretic scaling function $B$ that rescales the neutral
%pairwise coalescent times the appropriate amount to

%Strictly speaking, there is no way to rescale the coalescent under general
%background selection process, as genealogies deviate from the Kingman
%coalescent slightly under strong selection
%\parencite{Walczak2012-fi,Cvijovic2018-vd} and greatly under weak selection
%\parencite{OFallon2010-my}

%First, we approximate the coalescent effective population size, which
%determines expected pairwise diversity, from the variance effective population
%size. Then, we use a general quantitative genetic model of linked selection to
%find the variance effective population size under background selection. Since
%the impact of selection under this model is parameterized by the additive
%genetic fitness variance, we use an approximation for this based on the
%additive \emph{genic} variance, which depends on the rate of the ratchet.
%Previous work has been unable to find a simple closed-form solution to the rate
%of the ratchet because it depends on moments of the fitness distribution (e.g.
%fitness variation), which in turn depend on even higher-moments of the fitness
%distribution (e.g. the skew of the fitness distribution). While the ratchet
%should scale with the probability of fixation, the probability of fixation
%(Kimura, Malecot, XXX) is not accurate when multiple deleterious mutations are
%segregating simultaneously and in close linkage with one another. A key insight
%of \textcite{Santiago2016-mu} was that the classic equation for the probability
%of fixation is fairly robust as long as the correct effective population size
%is used. Their approach is to solve a system of two non-linear equations: one
%for the variance effective population size (which depends on the rate of the
%ratchet), and another for the time to fixation (which depends on the effective
%population size).

%Below, we step through the theory we use in our statistical methods. We
%re-derive central parts to illustrate where approximations have been made,
%which allows us to understand how these approximations may impact our
%statistical methods and inspect model fit in regions where these approximations
%are less likely to hold.

%\subsection*{Pairwise Diversity under Background Selection}

%The central quantity of interest for our statistical inference methods is
%average pairwise diversity along the chromosome. While other summaries of the
%data, such as the allele frequency spectrum, may be more informative about the
%evolutionary processes acting in a region, we lack population genetic theory
%for these quantities under most selective regimes and they require more degrees
%of freedom to characterize. By contrast, neutral diversity under an
%infinite-sites model is determined only by the per-basepair mutation rate $\mu$
%and the expected pairwise coalescent time $\E[T_2]$,

%\begin{align}
%  \E(\pi) = 2 \mu \E(T_2)
%\end{align}

%Let us assume that there is a constant \emph{drift-effective} population size
%$N_d$ which determines the rate of coalescence due only to neutral demographic
%processes. Additionally, we assume some background selection process that in
%expectation scales the effective population size by $b(t)$ at time $t$ (though
%this could in theory be due to any process). Then, the expected pairwise
%coalescent time is

%\begin{align}
%  \label{eq:T_2}
%  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(i) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
%\end{align}

%The time-dependency of $b(t)$ is required because the background selection
%process will usually alter the rate of coalesce through time. For example,
%under weak and intermediate selection, coalesce rates speed up as lineages move
%in the past under intermediate to weak selection, similar to population growth
%\parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.

%% Intuitively, this is
%% because lineages are unlikely to coalesce with numerous less fit lineages they
%% encounter as they go back in time, until ultimately coalescing in one of the
%% least-loaded fitness classes.

%This formulation is general and classic background selection theory (in the
%no-ratchet domain) is a special case of it. Under classic BGS and load theory,
%two randomly sampled lineages will each carry a number of deleterious mutations
%sampled from a Poisson stationary distribution. Because any lineage carrying $j
%> 0$ strongly deleterious mutations is destined to extinction, these lineages
%can only coalesce in a common ancestor in the least-loaded class with zero
%deleterious mutations. Under the Poisson stationary distribution, the size of
%this class is $f_0 N$, where $f_0 = e^{-\nicefrac{U}{s}}$, so the coalescent
%rate is scaled by $\nicefrac{1}{f_0}$. Once the lineages reach this fitness
%class, there are no fitness differences among individuals, so they coalesce
%following the Kingman process. A lineage moves backwards in time $t_d$
%generations through the ``delay phase" to the least-loaded class. Classic BGS
%theory ignores this delay phase, since $t_d \sim \log(j+1)/s << f_0 N$
%(Durrett, p. 213).

%%Here, we model the background selection process forward in time. We approximate
%%the expected pairwise diversity with the probability of heterozygosity,
%%$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
%%mating, such that

%%\begin{align}
%%  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
%%  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
%%              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
%%\end{align}

%%in the limit where $\mu \to 0$. This is equation 4 of
%%\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
%%population size.

%\subsection*{Forward in Time Models of Linked Selection}

%Next, show how quantitative genetic models of linked selection
%\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
%find expressions for $b(t)$ with by modeling the selective process forward in
%time. Their approach is to consider how the variance in neutral allele
%frequency change is impacted by a linked selection process.

%\subsubsection*{Variance in Neutral Allele Frequency in a Neutral System}

%For comparison, we start with the recursion for the variance in allele
%frequency change in a neutral Wright--Fisher model. Under binomial sampling of
%alleles, the random frequency $x_t$ is $2N x_t \sim \mathrm{Binom}(x_{t-1},
%2N)$. If we look forward in time on generation and treat $x_0$ as fixed,

%\begin{align}
%  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
%\end{align}

%and

%\begin{align}
%  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
%          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
%          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
%          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
%          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
%          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
%\end{align}

%Note that the cross-moment $\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right]
%= 0$ under neutral evolution. In general, the variance in allele frequency
%after $t$ generations of neutral evolution is,

%\begin{align}
%  \label{eq:var_freq}
%  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
%  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
%\end{align}

%This geometric series converges to the familiar expression,

%\begin{align}
%  \V(x_t) &= x_0(1-x_0)\left(1 - \left(1 - \frac{1}{2N}\right)^t \right).
%\end{align}

%which in the limit $t \to \infty$, converges to the Bernoulli variance
%$x_0(1-x_0)$ as the frequency either goes to 1 with probability $x_0$ or 0 with
%probability $1-x_0$.

%% Now, if the effective population size were to change each generation, equation
%% \eqref{eq:var_freq} would be,

%% \begin{align}
%%   \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
%% \end{align}

%% Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.

%\subsubsection*{Variance in Neutral Allele Frequency with Linked Selection}

%Note that equation XXX is a special case, when the evolutionary process has
%independent allele frequency changes through the generations and thus the
%cross-moment is zero. Selection generates correlations between allele frequency
%changes at nearby neutral sites. These correlations created by linked selection
%inflate the variance in neutral allele frequency $\V(x_t)$ beyond what it would
%be in a system without linked selection operating. The magnitude of this
%inflation term is set by the selection process and its dynamics, linkage
%disequilibrium, and the recombination rate. Even k

%The theory of this began with \textcite{Robertson1961-ho},
%and was thoroughly worked out by Santiago and Caballero
%(\citeyear{Santiago1995-hx}, \citeyear{Santiago1998-bs}; see also
%\cite{Barton2000-zg} and \cite{Buffalo2019-qs}). Our BGS statistical methods
%employ the extension of this model for weak BGS \parencite{Santiago2016-mu}.
%For completeness, we step through a quick derivation of the linked selection
%inflation factor models here, based on
%\textcite{Santiago1995-hx,Santiago1998-bs}.

%Quantitative genetic models of linked selection

%%%%%%

%One way we can model these dependencies is to partition $\Delta_i$ into drift
%and selection components. Over evolutionary replicates, the drift components
%are independent and the selection components have some degree of
%autocorrelation. Let us partition $\Delta x_i = \Delta_S x_i + \Delta_D x_i$,

%\begin{align}
%  \V(x_2) &= \V\left(\Delta x_2 + \Delta x_1\right) \\
%          &= \V(\Delta_D x_2) + \V(\Delta_D x_1) + 2 \cov(\Delta_S x_2, \Delta_S x_1) \\
%          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\Delta_S x_2, \Delta_S x_1) 
%\end{align}

%since $\Delta_D x_i \perp \Delta_D x_j$ for $i \ne j$. The covariances are
%determined by the selective process dynamics, linkage disequilibrium, and the
%recombination rate. These covariances are throughly worked out in
%\textcite{Santiago1995-hx} and \textcite{Buffalo2019-qs}, so we give a short
%derivation here. Throughout, we assume random mating, hermaphroditic
%individuals, and a constant population size. We will use angular brackets to
%indicate an average over all diploid individuals in the population, of some
%value with an individual-level index. For example, if we denote the number of
%allele an individual carries in generation $t$ as $z_{t,i} \in \{0, 1, 2\}$,
%the population frequency at time $x_t$ is

%\begin{align}
%  x_t = \langle z_t \rangle = \frac{1}{2N} \sum_{i=1}^N z_{t,i}
%\end{align}

%Across evolutionary replicates, each parent $i$ passes one of their two neutral
%alleles to a random number $k_i$ offspring. The \emph{heritable fitness} of an
%individual $f_i$ is defined as the expected number of offspring passed down
%across evolutionary replicates, such that $k_i = f_i + \varepsilon_i$.  Some
%selection process determines the distribution of $f_i$, but the this occurs
%independently of the process that sorts neutral alleles into diploid
%individuals. We are unsure of the neutral allele count in an individual in the
%initial generation, $z_{t,0}$, so we condition on it and take expectations
%later. Then, the frequency in the initial and first generations are,

%\begin{align}
%  x_0 = \langle z_0 \rangle = \frac{1}{2N} \sum_{i=1}^N z_{0,i} \\
%  x_1 = \langle z_1 \rangle = \frac{1}{2N} \sum_{i=1}^N \left( k_i \frac{z_{0,i}}{2} + \sum_{j=1}^{k_i} \delta_{i,j} \right)
%\end{align}

%The first term is the average number of neutral alleles each parent passes to
%its $k_i$ offspring. In reality, alleles are discrete, so each heterozygous
%parent also passes a random $\pm \nicefrac{1}{2}$ to each offspring, which
%leads to the second term. The variable $\delta_{i,j}$ is 0 if the parent is a
%homozygote, but if the parent is a heterozygote, is $\pm \nicefrac{1}{2}$ with
%equal probability over evolutionary replicates.

%Since individuals are exchangeable with their contribution to the
%heterozygosity term and population size is constant,

%\begin{align}
%  \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{i,j} = \sum_{i=1}^N \delta_{i} 
%\end{align}

%We can further partition the offspring number into heritable and non-heritable
%parts,

%\begin{align}
%  x_1 &= \frac{1}{2N} \sum_{i=1}^N \left( (f_i + \varepsilon_i) \frac{z_{0,i}}{2} + \delta_{i} \right) \\
%      &= \frac{1}{4N} \sum_{i=1}^N f_i z_{0,i} + \frac{1}{4N} \sum_{i=1}^N \varepsilon_i z_{0,i} + \frac{1}{2N} \sum_{i=1}^N \delta_{i} \\
%      &=  \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle \\
%\end{align}

%% TODO factor 1/2 error

%\begin{align}
%  x_1 - x_0 &= \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle - \langle z_0 \rangle \\
%            &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
%\end{align}

%Now, we look at the variance of this over evolutionary replicates. By symmetry,
%the expected change in a neutral allele is $\E(\Delta x_t) = 0$ since we track
%an arbitrary neutral allele. Then the variance $\V(\Delta x_1) = \E( {\Delta
%x_1}^2)$,

%\begin{align}
%  \V(\Delta x_1) &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
%\end{align}

%OLD
%Let us imagine there is some stochastic process $b(t)$ that rescales the coalescent
%rate,

%\begin{align}
%  \V(x_t) &= \frac{x_0(1-x_0)}{2N(0)}\left(\sum_{i=1}^t \prod_{j=1}^{i-1} \left(1 - \frac{1}{2b(j)N}\right) \right)
%\end{align}

%\begin{align}
%  \var(p_3 - p_0) = \E & \left( \bigg( \underbrace{S_1(1 + (1-r) G + (1-r)^2 G^2)}_{\text{creation and persistence of generation 1 associations} \;\; := \; S_1 Q_3} \right. +  D_1 + H_1+ \nonumber \\
%                       & \underbrace{S_2(1 + (1-r) G)}_{\text{creation and persistence of generation 2 associations} \;\; := \; S_2 Q_2}+ D_2 + H_2 + \nonumber \\
%                       &  \underbrace{S_3}_{\text{creation of generation 3 associations} \;\; := \; S_3 Q_1  }\left. + D_3 + H_3 \bigg)^2 \right). &
%    \label{eq:scr-var2}
%\end{align}
%%

%Note the similarity between the equations.

%\begin{align}
%  V &\propto \frac{1}{2b(0)N_d}\left(\sum_{i=1}^\infty \prod_{t=1}^{i-1} \left(1 - \frac{1}{2b(t)N_d}\right) \right) \\
%  W &\propto \sum_{i=1}^\infty \frac{i}{2 b(i) N_d} \prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right) \\
%\end{align}

%Let us define the function $a(t)$ to be the product, which is the cumulative
%failure to coalesce up to time $t$.

%\begin{align}
%  a(t) = \prod_{i=1}^{t-1} \left( 1-\frac{1}{2 b(i) N_d} \right)
%\end{align}

%Comparing this to equation \eqref{eq:T_2} up to constant factors,

%\begin{align}
%  V &\propto \sum_{i=1}^\infty a(t) \\
%  W &\propto \sum_{i=1}^\infty i \; a(t) \\
%\end{align}

%end{align}

%XXX we swap indices

%Now, if we assume that the final coalesce in the waiting time process were to
%have a fixed rate of $\nicefrac{1}{2b(0)N_d}$, this

%we see that both the waiting time process and the variance process

%Intuitively, these define a stochastic selection process in two ways through
%the function $b(t)$ that modulates the rate of pairwise coalesce.

%where if $N_e(t) = N_e$ and $t=1$, this is the familiar Wright--Fisher variance
%$\V(p_1) = p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency
%changes per generation are independent, and neutral diversity is set by this
%single $N_e$.

%XXX

%Since we track an arbitrary neutral allele's frequency $p_i$, by symmetry,
%linked selection has no impact on the direction of change, $\E(\Delta p_i) =
%0$. We will work through a short proof of $\V(p_t)$ by induction following the
%approach of \textcite{Santiago1995-hx}. Throughout, we will use angular
%brackets to indicate an average over all diploid individuals in the population,
%of some value with an individual-level index. For example, if we denote the
%number of allele an individual carries in generation $t$ as $x_{t,i} \in \{0,
%1, 2\}$, the population frequency at time $p_t$ is

%\begin{align}
%  p_t = \langle x_t \rangle = \frac{1}{N} \sum_{i=1}^N x_{t,i}
%\end{align}

%\begin{align}
%  \Delta_H p &= \langle x (f-1) \rangle \\
%  \Delta_N p &= \langle x d \rangle \\
%  \Delta_M p &= \langle \Delta \rangle \\
%\end{align}

%Assuming uninherited, random environments, only the $\Delta_{_H} p$ terms
%create autocorrelation in $p_t$. The autocorrelation is created because during
%the random pairing of neutral alleles to fitness backgrounds, associations are
%created between an individual $i$'s allele frequency and their fitness $f_i$.
%This is because both fitness and the neutral alleles reside in the same
%individual in the present generation.

%We can imagine now how these associations change over evolutionary replicates
%in the next generation. In a fraction $r$ of the cases, the neutral allele 
%recombines off (or segregates away, if they are on different chromosomes) the
%fitness background. In a faction of $1-r$ XXX

%\begin{align}
%  \cov(\Delta_H p_t, \Delta_H p_s) &= \E(\langle x_t (f_t-1) \rangle , \langle x_s (f_s-1) \rangle)
%\end{align}

%Note that this covariance is only

%In the first generation, the frequency change
%can be partitioned into three components. Conditioning on the parent's
%heritable fitnesses, we can imagine constructing an infinite number of
%evolutionary replicates. Across all replicates, a random neutral allele's is
%perturbed

%\begin{align}
%  p_1 - p_0 = \sum_{i=1}^{N} 
%\end{align}

%Since the variance in neutral allele frequency change is also determined by the
%pairwise coalescence rate, this provides us with a way to model $b(t)$ under
%linked selection.

%These models derive an expression for the variance-effective population size of
%a focal neutral site, $N_v(t)$, based on how linked fitness variation perturbs
%the frequency trajectory at this site.  The variance-effective population size
%for a constant $N_e$ is defined by the variance in allele frequency change at
%the neutral site, over evolutionary replicates,

%where if $t=1$, this is the familiar Wright--Fisher variance $\V(p_1) =
%p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency changes per
%generation are independent, and neutral diversity is set by this
%single $N_e$.

%However when selection is acting at linked sites, neutral diversity is
%determined by the long-run $\V(p_t)$ as $t \to \infty$. Intuitively, this is
%because the rate of drift is magnified by autocorrelation in allele frequency
%change (XXX), and diversity is determined by the cumulative impact of these
%autocorrelations in the long-run. This autocorrelation effectively reduces the
%coalescence rate in a time-dependent manner, analogous to equation
%\eqref{eq:T_2}. Thus, a single constant $N_e$ is insufficient to describe
%levels of neutral variation, since it is insufficient

%The variance in allele frequency change after $t$ generations is

%\begin{align}
%  \V(p_t) = p_0(1-p_0)\left(1 - \prod_{i=1}^t \left(1-\frac{1}{2N_e(t)}\right)\right)
%\end{align}

%This is the basis of the quantitative genetic models of linked selection: a
%forward in time model for $\V(p_t)$ implies a pairwise coalescence rate
%function $b(t)$, which determines statistics based on pairwise coalescence such
%as $\E(\pi)$.

%Next, we consider how fitness variation impacts $\V(p_t)$.

%Throughout, we will take the variance $\V(p_t)$ to be the variance over
%evolutionary replicates. Since the variance is for a random neutral allele's
%frequency,

%a reasonable approximation to how background selection process
%rescales the coalescent rate at this site is given by $b(t) =
%\nicefrac{N_v(t)}{N_d}$.

%\begin{align}
%  N_v = \lim_{t \to \infty} \frac{p_0 (1 - p_0) - \var(p_{t-1})}{2\var(p_t) - \var(p_{t-1})}
%\end{align}

%\subsection*{Linked Selection Models for Variance-Effective Population Size}

%We wish to derive an expression for the reduction factor $B$. Ideally, this
%reduction factor would rescale the coalescent-effective population size, but
%this would require knowing the coalescent rate function $\lambda(t)$ over the
%evolutionary history $0 < t < \infty$. Instead, following previous work
%\parencite{Santiago1995-hx,Santiago1998-bs}, we derive the rescaling function
%$B$ relative to the variance-effective population size.

%Following past quantitative genetic models of linked selection
%\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
%we consider the long-run variance effective population size,

%\begin{align}
%  N_v := \lim_{t \to \infty} \frac{p_0(1-p_0)}{2\var\left(p_t - p_0\right)}
%\end{align}

%% ADD TO INTRO TODO
%Comment about characteristic scale $\nicefrac{\mu}{s}$. Is the fact that we see
%BGS on megabase scales because this is the only scale we can detect strong
%selection under?

>>>>

Finally, we also used simulations to test the sensitivity of our method to
violations of two assumptions: constant population size and recessivity of
selective effects. Since we find little difference in parameter estimates among
YRI and bottlenecked out-of-Africa (CEU and CHB) samples, we tested whether a
population expansion would bias our method. We expansion rates like those
estimated for Europeans by \textcite{Gutenkunst2009-pg} had little effect on
parameter estimates (Supplementary Section \ref{supp:sim-assum}). By contrast,
we find our estimates are downwardly biased when mutational effects are weak
and recessive, but are accurate under recessive strong selection. This occurs
since the fixation probabilities of recessive mutations are in reality much
lower than they are under our model assuming additivity.

\subsection*{Predicted Diversity and Residual Signal}

@@

While we have previously shown that out-sample predicted diversity matches
observed diversity (Section XXX), we also investigated whether there were
regions of the genome that were particularly poor-fitting. Comparing predicted
against observed diversity along chromosomes (see Supplementary Materials
Section \ref{supp:chrom-fit}) we overall find close correspondence consistent
with the high $R_\text{LOCO}^2$. However, we find a few large (tens of
megabases) regions with systematically poorer fit (Supplementary Materials
Figure \ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one
such region from XXX to XXX. Interestingly, predicted diversity closely follows
the peaks and troughs of this region. We note that a smaller segment within
this region has been found by a genome-wide scan for associative overdominance
\parencite{Gilbert2020-aw}.  For model XXX

Another model diagnostic is to inspect whether observations are systematically
different from predictions. We confirm the finding of \textcite{Murphy2022-sj}
that regions predicted to experience little reduction in diversity due to
background selection have higher diversity than expected (Figure
\ref{fig:figure-5}B, orange line). Their work suggested that this could reflect
ancient introgression between archaic humans and ancestors of contemporary
humans. Finally, the variance around observed and predicted diversity levels
falls very close to what we would expect under the theoretic
coalescence-noise-only expectation ($R_\text{coal}^2$).

\subsection*{Predicted Substitution Rates, and Fitness Variance}

@@

Since our B' method is based upon quantitative genetic models of linked
selection that explicitly incorporates the substitution process, fitting our
model to genomic data leads to predictions that can be explicitly tested. In
particular, there are two key predicted quantities: (1) the substitution rates
per segment, and (3) the fitness variance per segment.

@@

@@
